---
title: "Shiny App Presentation"
author: "Plato Karageorgis"
date: '2022-06-23'
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Introduction

This is a presentation for the shiny app on this link -->  https://platon.shinyapps.io/Shiny/


## Prerequisites

To run this app you need to:

  - Install shiny package with install.packages("Shiny") and load the library with library(shiny)
  - Install C50 package with install.packages("C50") and load the library with library(C50)

## Documentation

The documentation of the app can be found on the link below

  - https://github.com/Platon-7/Platon-7.github.io/blob/ghpages/Developing%20Data%20Products/Week%204/Shiny%20App/Documentation.rmd
  
## A quick look to the data

```{r, echo = TRUE}
  summary(iris)
```
## A plot of our decision tree

```{r, echo = FALSE}
  options(warn=-1)
  library(C50)
  inner <- iris[,1:4]
  outsider <- iris[,5]
  model <- C5.0(inner, outsider, control = C5.0Control(noGlobalPruning = FALSE))
  plot(model, main="C5.0 Decision Tree - Pruned")
```

